Artin Group Presentations Arising from Cluster Algebras

### 2014-07-29 Paper.tex
emily edits to section 1/2

\begin{abstract}
In 2003, Fomin and Zelevinsky proved that finite-type cluster algebras can be classified by Dynkin diagrams. Then in 2013, Barot and Marsh defined the presentation of a reflection group associated to a Dynkin diagram in terms of an edge-weighted, oriented graph, and proved that this group is invariant (up to isomorphism) under diagram mutations. In this paper, we extend Barot and Marsh's results to Artin group presentations, defining new generator relations and showing mutation-invariance for these presentations.
\end{abstract}

@@

In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras in order to make further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite-type} cluster algebras, that is, cluster algebras whose variables are generated through mutation on a finite number of \textit{seeds}. In the sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite-type cluster algebras can be classified by Dynkin diagrams. \\
\indent Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite-type cluster algebra. They also proved that this group presentation is invariant up to isomorphism under the mutation equivalence relation.That is, given a diagram $\Gamma$ and a diagram mutation equivalent to $\Gamma$, denoted $\Gamma^{\prime} = \mu_k(\Gamma)$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group representations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively.

@@

A \textit{cluster algebra} is an integral domain which can be generated by a set of elements called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{intial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \{x_1, \ldots, x_n\}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetric matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{-B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\boldsymbol{x}, B) = (\textbf{x}^{\prime}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} - B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
																B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
									\end{cases}
\end{displaymath}
and $\textbf{x}^{\prime} = {x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n}$. 
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = {x_1, \ldots, x_n}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

 A cluster algebra is said to be of \textit{finite type} if the number of cluster variables that generate it is finite (if it has finitely many seeds). For each finite type cluster algebra, we can associate to its corresponding skew-symmetrisable matrix an edge-weighted, oriented graph, called a \textit{diagram}. We will often denote this diagram by $\Gamma$, and the vertex set of $\Gamma$ by $\V$. We will denote two connected vertices by $i \rightarrow j$, or by $i\--\ j$ if the orientation is not specified. The diagram is determined by, for $i, j \in \V$, $i \xrightarrow{w} j$ if and only if $B_{ij} > 0$ and $w = |B_{ij}B_{ji}|$ is the weight of the edge. A skew-symmetrisable matrix $B$ is \textit{2-finite} if $|B_{ij}B_{ji}| \leq 3$ for $i, j \in {1, \ldots, n}$.  By \cite[7.5]{FZ02}, we have that if $B$ is 2-finite, all 3-cycles in the unoriented graph underlying our diagram must be oriented cyclically. \\

Just as we can define mutations of the seeds of a cluster variable, we can also define mutations of a diagram associated to a cluster algebra of finite type by the following set of rules:
\begin{prop}\cite[Proposition 1.4]{BM13}
Let $B$ be a 2-finite skew-symmetrisable matrix. Then $\Gamma(\mu_k(B))$ is uniquely determined by $\Gamma(B)$ as follows: 
\begin{itemize}
\item Reverse the orientations of all edges in $\Gamma(B)$ incident with $k$ (leaving the weights unchanged)
\item For any path in $\Gamma(B)$ of form $i \xrightarrow{a} k \xrightarrow{b} j$ (i.e. with $a,b$ positive), let $c$ be the weight on the edge $j \rightarrow i$, taken to be zero if there is no such arrow. Let $c'$ be determined by $c'\geq 0$ and 
$c+c' =$ max$(a,b)$. 
Then $\Gamma(B)$ changes as in Figure \ref{mutation}, taking the case $c' = 0$ to mean no arrow between $i$ and $j$.
\end{itemize}
\end{prop}

@@

Then, they define $W(\Gamma)$ to be the group generated by $s_i$, for $i \in \V$, under the following relations. Note that $e$ will denote the identity element of $W(\Gamma)$. 
\begin{enumerate}
\item $s_i^2 = e$ for all $i$;
\item $(s_is_j)^{m_{ij}} = e$ for all $i \neq j$;
\item For any chordless cycle $C$ in $\Gamma$, where
\begin{displaymath}
C = i_0 \rightarrow i_1 \rightarrow \cdots \rightarrow i_{d-1} \rightarrow i_0
\end{displaymath}
and all of the weights $w_k$ are $1$ or $w_0 = 2$, we have
\begin{displaymath}
(s_{i_0}s_{i_1}\cdots s_{i_{d-2}}s_{i_{d-1}}s_{i_{d-2}}\cdots s_{i_1})^2 =e.
\end{displaymath}
\todo{Maybe it's worth presenting definition 2.4 before this, and then saying it is equivalent to removing certain the (R3)(b) relations. Don't they actually present the definition you refer to in Definition 2.4 as the presentation? You say it is an alteration there, but I think it's the original definition}
\end{enumerate}

@@

The rest of the paper will be devoted to building up analogous relations, defined in ~\ref{grp def} to prove a similar result in the case of Artin groups. For $\Gamma$ a diagram of finite type, we define the associated Artin Group as follows.

\begin{defn}
The associated Artin group $A_\Gamma$ is generated by $s_i,$ where there is one $s_i$ for each vertex $i$ in $\Gamma.$ These generators are subject to the following relations
\begin{itemize}
	\item[(R2)] With $m_{ij}$ as defined in ~\ref{mij_defn}, for all $i \neq j,$ we add the relations
	$\langle s_i,s_j \rangle^{m_{ij}}= \langle s_j,s_i \rangle^{m_{ij}}.$

\item[(R3)] For every ordered chordless cycle of the form

\begin{tikzcd}
i_0 \arrow{r} & i_1 \arrow{r} & \cdots \arrow{r} & i_{d-1} \arrow{r} & i_0
\end{tikzcd}

such that either
\begin{enumerate}
	\item All edges in the chordless cycle are of weight 1 or 2 and the edge $i_{d-1}\rightarrow i_0$ has weight 2.
	\item All edges in the chordless cycle have weight $1.$
\end{enumerate}
we include the relation
\begin{align*}
	s_{a}s_{a+1}^{-1}s_{a+2}^{-1}\dots s_{a-2}^{-1}s_{a-1}s_{a-2}s_{a-3}\dots s_{a+1} = s_{a+1}^{-1}\dots s_{a-3}^{-1}s_{a-2}^{-1}s_{a-1}s_{a-2}\dots s_{a+1}s_{a},
\end{align*}
where subscripts are taken $\pmod d.$ We will denote this relation by r(a, a+1). 
\end{itemize}
\end{defn}

 We will use these group relations to prove the mutation invariance of $A_{\Gamma}$ in Section \ref{sec:proof_of_main}

@@

\begin{defn}
A {\it chordless cycle} of an unoriented graph $G$ is a connected subgraph $H \subset G$ such that the number of vertices in $H$ is equal to the number of edges in $H,$ and the edges in $H$ form a single cycle.  
\end{defn}

>>>>

\begin{abstract}
In 2003, Fomin and Zelevinsky proved that finite type cluster algebras can be classified by Dynkin diagrams. Then in 2013, Barot and Marsh defined the presentation of a reflection group associated to a Dynkin diagram in terms of an edge-weighted, oriented graph, and proved that this group is invariant (up to isomorphism) under diagram mutations. In this paper, we extend Barot and Marsh's results to Artin group presentations, defining new generator relations and showing mutation-invariance for these presentations.
\end{abstract}

@@

In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras in order to make further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite type} cluster algebras, that is, cluster algebras whose variables are generated through mutation on a finite number of \textit{seeds}. In the sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite type cluster algebras can be classified by Dynkin diagrams. \\
\indent Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite type cluster algebra. They also proved that this group presentation is invariant up to isomorphism under the mutation equivalence relation.That is, given a diagram $\Gamma$ and a diagram mutation equivalent to $\Gamma$, denoted $\Gamma^{\prime} = \mu_k(\Gamma)$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group representations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively.

@@

A \textit{cluster algebra} is an integral domain which can be generated by a set of elements called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{intial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \{x_1, \ldots, x_n\}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetric matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{-B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\boldsymbol{x}, B) = (\textbf{x}^{\prime}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} - B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
																B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
									\end{cases}
\end{displaymath}
and $\textbf{x}^{\prime} = \left{ x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n \right}$. 
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = \left{ x_1, \ldots, x_n \right}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

 A cluster algebra is said to be of \textit{finite type} if the number of cluster variables that generate it is finite (if it has finitely many seeds). For each finite type cluster algebra, we can associate to its corresponding skew-symmetrisable matrix an edge-weighted, oriented graph, called a \textit{diagram}. We will often denote this diagram by $\Gamma$, and the vertex set of $\Gamma$ by $\V$. We will denote two connected vertices by $i \rightarrow j$, or by $i\--\ j$ if the orientation is not specified. The diagram is determined by, for $i, j \in \V$, $i \xrightarrow{w} j$ if and only if $B_{ij} > 0$ and $w = |B_{ij}B_{ji}|$ is the weight of the edge. A skew-symmetrisable matrix $B$ is \textit{2-finite} if $|B_{ij}B_{ji}| \leq 3$ for $i, j \in \left{ 1, \ldots, n \right}$.  By \cite[7.5]{FZ02}, we have that if $B$ is 2-finite, all 3-cycles in the unoriented graph underlying our diagram must be oriented cyclically. \\

Just as we can define mutations of the seeds of a cluster variable, we can also define mutations of a diagram associated to a cluster algebra of finite type by the following set of rules:
\begin{prop}\cite[Proposition 1.4]{BM13}
Let $B$ be a 2-finite skew-symmetrisable matrix. Then $\Gamma(\mu_k(B))$ is uniquely determined by $\Gamma(B)$ as follows: 
\begin{itemize}
\item Reverse the orientations of all edges in $\Gamma(B)$ incident to $k$ (leaving the weights unchanged)
\item For any path in $\Gamma(B)$ of form $i \xrightarrow{a} k \xrightarrow{b} j$ (i.e. with $a,b$ positive), let $c$ be the weight on the edge $j \rightarrow i$, taken to be zero if there is no such arrow. Let $c'$ be determined by $c'\geq 0$ and 
$c+c' =$ max$(a,b)$. 
Then $\Gamma(B)$ changes as in Figure \ref{mutation}, taking the case $c' = 0$ to mean no arrow between $i$ and $j$.
\end{itemize}
\end{prop}

@@

Then, they define $W(\Gamma)$ to be the group generated by $s_i$, for $i \in \V$, under the following relations. Note that $e$ will denote the identity element of $W(\Gamma)$. 
\begin{enumerate}
\item $s_i^2 = e$ for all $i$;
\item $(s_is_j)^{m_{ij}} = e$ for all $i \neq j$;
\item For any chordless cycle (as defined in Definition \ref{chordlesscycle} $C$ in $\Gamma$, where
\begin{displaymath}
C = i_0 \xrightarrow{w_1} i_1 \xrightarrow{w_2} \cdots \xrightarrow{w_{d-1}} i_{d-1} \xrightarrow{w_0} i_0
\end{displaymath}
and all of the weights $w_k$ are $1$ or $w_0 = 2$, we have
\begin{displaymath}
(s_{i_0}s_{i_1}\cdots s_{i_{d-2}}s_{i_{d-1}}s_{i_{d-2}}\cdots s_{i_1})^2 =e.
\end{displaymath}
%\todo{Maybe it's worth presenting definition 2.4 before this, and then saying it is equivalent to removing certain the (R3)(b) relations. Don't they actually present the definition you refer to in Definition 2.4 as the presentation? You say it is an alteration there, but I think it's the original definition}
\end{enumerate}

@@

The rest of the paper will be devoted to building up analogous relations, defined in ~\ref{grp def} to prove a similar result in the case of Artin groups. For $\Gamma$ a diagram of finite type, we define the Artin group associated to $\Gamma$ as in Section \ref{sec:defn_artingroup}. We will then use the group relations presented in this definition to prove the mutation invariance of $A_{\Gamma}$ in Section \ref{sec:proof_of_main}.

@@

\begin{defn}
\label{chordlesscycle}
A {\it chordless cycle} of an unoriented graph $G$ is a connected subgraph $H \subset G$ such that the number of vertices in $H$ is equal to the number of edges in $H,$ and the edges in $H$ form a single cycle.  
\end{defn}
